A Web-Based Data Visualization Platform for MATSim

### 2020-05-15 src/paper.tex
add emissions figure

@@

\begin{abstract}
There are many tools available for analyzing MATSim transport
simulation results, both open-source and commercial. This research
builds a new open-source visualization platform for MATSim outputs that
is entirely web-based. After initial experiments with many different web
technologies, a client/server platform design emerges which leverages
the advanced user interface capabilities of modern browsers on the
front-end, and relies on back-end server processing for more CPU-intensive
tasks. The initial platform is now operational and includes several
aggregate-level visualizations including origin/destination flows,
transit supply, and emissions levels; as well as a fully disaggregate
traffic animation visualization. These visualizations are general
enough to be useful for various projects. Further work is needed
to make them more compelling and the platform more useful for practitioners.
\end{abstract}

@@

The rapid advancement of Internet browsing technologies in the last
five years has enabled the web browser to do things much more
"application"-like than ever before: background processing,
three-dimensional rendering using GPU acceleration, offline
support, and more. The combination of these technologies and their
standard implementations on every popular hardware type and operating system now makes the web a very compelling platform.

@@

Currently, analysis of MATSim outputs ends up in research reports, PDFs, video screen-recordings, and presentations. An online dashboard of results which a user could explore and manipulate would not only be more interactive, but might also reveal findings that the original analysts hadn't anticipated.

@@

\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page. CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
\item[(ii)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
\item[(iv)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's. Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
\item[(v)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow baiun truly multi-threaded code inside a browser.

@@

For this research, we are deliberately exploring the latest \textit{standard} web technologies, with the expectation that access to these technologies will become more and more common in the future. Thus, we are targeting the most recent versions of modern web browsers as of 2019, including Google Chrome, Mozilla Firefox,
and Apple Safari. All three browsers fully support the above-listed technologies, and importantly, all three auto-update automatically, ensuring that most users of those browsers stay current as these technologies evolve.

@@

No proprietary or closed licensing schemes were considered, primarily because excellent proprietary visualization packages for MATSim already exist. Creating a competing product would be duplicative and unnecessary, and would not further the research goal of determining whether web-based technology is now advanced enough to work with MATSim outputs. The goal of this research is not to replace existing, proprietary solutions, but rather to complement them.

@@

This matches the license of MATSim itself. Several other open-source licenses were considered, including the MIT License and the Apache Public License, but the benefit of sharing a common license with MATSim outweighs any perceived benefits of switching to other open licenses.

@@

MATSim network files are small enough to fit in RAM, but MATSim plan files and event files can be much larger than RAM, necessitating careful consideration about how to handle them.

@@

The primary framework used to build the application is known as "Vue JS" (Vue, available at vuejs.org). Vue is a framework for building interactive user interfaces on the web; essentially it provides the glue between the items a user clicks and the code that runs when they do so. Vue provides many services which allow a web page to behave more like a full-featured application, including state management, routing between different URLs, and componentization of code in a way that encourages code reuse and loose coupling. Vue depends on Javascript, which means it does not work on for users who have disabled scripting in their browser.

Vue is most often used to write so-called "single page applications" which are websites that behave like desktop applications. Most large, popular websites such as Github, Twitter, and Facebook all employ this paradigm, meaning the site handles page transitions and URL's as if they are all in one common namespace, and the user thinks less about visiting URLs and more about navigation through visual components to accomplish tasks. This matches our use case.

@@

\item[(i)] Origin/destination flows between aggregate areas, so called "spider diagrams" (Figure~\ref{Fod})
\item[(i)] Link flows, by time of day and mode (Figure~\ref{Flinkvols})
\item[(iii)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links. (Figure~\ref{Ftransit})
\item[(iv)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (Figure~\ref{Fsankey})
\item[(v)] Emissions levels on a geographic hexagonal grid basis

@@

@@

These usability problems were eventually solved by surveying other technical websites which organize and present large amounts of data. One website in particular, \url{Github.com}, was found to be well-liked by test users and has a similar hierarchical view of data: users can belong to organizations, and both organizations and users can create projects ("repositories" on Github) which may large numbers of files.

By adopting a file, project, and user paradigm similar to that employed by Github, users were immediately more familiar and had less to learn. The current site adheres to this so closely that test users often refer to the platform as "MatHub".

@@

\textbf{Map layer limitations.} The Mapbox GL mapping library allows lines, points, and polygons to be displayed on top of a base map. There seems to be a number of map elements beyond which performance becomes very slow; various techniques can be used such as aggregating elements at further-out zoom levels to get around this. Another option is to only use Mapbox GL for the base map, and to use the WebGL graphics libraries to draw large numbers of elements directly on top of the map. This is the technique used for the traffic animation, which can easily support tens of thousands of elements (all in motion!) simultaneously. For additional visualizations which have large numbers of grahical elements, more research will need to be done on layering WebGL elements on top of the base map.

@@

\bibitem[Strippgen, David (2016)]{R2}
Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 225–234.

\bibitem[Rieser, Marcel (2016)]{R3}
Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–224.

>>>>

% -- to allow highlight changed sections in yellow:
\usepackage{color,soul}

@@

\begin{abstract}
There are many tools available for analyzing MATSim transport
simulation results, both open-source and commercial. This research
builds a new open-source visualization platform for MATSim outputs that
is entirely web-based. After initial experiments with many different web
technologies, a client/server platform design emerges which leverages
the advanced user interface capabilities of modern browsers on the
front-end, and relies on back-end server processing for more processor-intensive
tasks. The initial platform is now operational and includes several
aggregate-level visualizations including origin/destination flows,
transit supply, and emissions levels; as well as a fully disaggregate
traffic animation visualization. These visualizations are general
enough to be useful for various projects. Further work is needed
to make them more compelling and the platform more useful for practitioners.
\end{abstract}

@@

The rapid advancement of Internet browsing technologies in the last
five years has enabled the web browser to do things much more
"application"-like than ever before: background processing,
three-dimensional rendering using \hl{graphics card (GPU)} acceleration, offline
support, and more. The combination of these technologies and their
standard implementations on every popular hardware type and operating system now makes the web a very compelling platform.

@@

Currently, analysis of MATSim outputs ends up in research reports, \hl{Adobe (PDF)} documents, video screen-recordings, and presentations. An online dashboard of results that a user could explore and manipulate would not only be more interactive, but might also reveal findings that the original analysts hadn't anticipated.

@@

\item[(i)] \textbf{HTML 5} improves and standardizes the "document model" of what constitutes a web page and how it is specified.
\item[(i)] \textbf{CSS 3} is a styling language that enables fine-grained layout and styling of individual elements on a page. CSS 3 defines in a consistent, standard way the details of elements including color, size, layout, and animation of page elements.
\item[(ii)] \textbf{WebGL} provides in-browser support for the 3D-accelerated graphics capabilities of modern machines.
\item[(iv)] \textbf{ECMAScript 6} is an updated (2015) version of the venerable Javascript scripting language that has been part of the web platform since the early 1990's. Recent versions of Javascript eliminate the more problematic aspects of the language and make it easier for developers to create bug-free, efficient code.
\item[(v)] \textbf{Web Workers} are a recent (2013) addition to the web platform that allow background processing for long-running tasks. Before Web Workers, there was no way to run truly multi-threaded code inside a browser.

@@

For this research, we are deliberately exploring the \textit{latest standard} web technologies, with the expectation that access to these technologies will become more and more common in the future. Thus, we are targeting the most recent versions of modern web browsers as of 2019, including Google Chrome, Mozilla Firefox,
and Apple Safari. All three browsers fully support the above-listed technologies, and importantly, all three auto-update automatically, ensuring that most users of those browsers stay current as these technologies evolve. \hl{This does leave out older browsers such as Internet Explorer and Edge, as well as browsers with small market share. Supporting those browsers would require much more deliberate and time-consuming testing, which while not impossible, is currently out of scope. For some government users this is acknowledged as possibly problematic, since use of Internet Explorer 11 is still mandated at some agencies.}

@@

No proprietary or closed licensing schemes were considered, primarily because excellent proprietary visualization packages for MATSim already exist. Creating a competing product would be duplicative and unnecessary, and would not further the research goal of determining whether web-based technology is now advanced enough to work with MATSim outputs. The goal of this research is not to replace existing proprietary solutions, but rather to complement them.

@@

This matches the license of MATSim itself. Several other open-source licenses were considered, including the MIT License and the Apache Public License, but the benefit of sharing a common license with MATSim outweighs any perceived benefits of switching to other open licenses \hl{The main benefit of those other open licenses is compatibility with commercial software. Given its narrow scope, this research is not likely to be of interest to existing commercial vendors. Furthermore, the research team does not intend to turn this into a commercial product. Thus the GPL license is a good fit.}

@@

MATSim network files are small enough to fit in \hl{computer memory (RAM)}, but MATSim plan files and event files can be much larger than RAM, necessitating careful consideration about how to handle them.

@@

The primary framework used to build the application is known as "VueJS" or Vue (available at \url{vuejs.org}). Vue is a framework for building interactive user interfaces on the web; it provides the glue between the items a user clicks and the code that runs when they do so. Vue provides many services which allow a web page to behave more like a full-featured application, including state management, routing between different URLs, and componentization of code in a way that encourages code reuse and loose coupling. Vue depends on Javascript, which means it \hl{does not work for users} who have disabled scripting in their browser.

Vue is most often used to write so-called "single page applications" which are websites that behave like desktop applications. Most large, popular websites such as Github, Twitter, and Facebook all employ this paradigm, meaning the site handles page transitions and URLs as if they are all in one common namespace, and the user thinks less about visiting URLs and more about navigation through visual components to accomplish tasks. This matches our use case.

@@

\item[(i)] Origin/destination flows between aggregate areas, so-called "spider diagrams" (Figure~\ref{Fod})
\item[(i)] Link flows, by time of day and mode (Figure~\ref{Flinkvols})
\item[(iii)] Transit supply explorer, which displays all transit routes and allows the user to see which routes serve specific stops and links. (Figure~\ref{Ftransit})
\item[(iv)] Sankey diagrams, which can be used to depict changes/flows between between scenarios across multiple choices, such as shifts in mode between two scenarios (Figure~\ref{Fsankey})
\item[(v)] Emissions levels on a geographic hexagonal grid basis (Figure~\ref{Femissions})

@@

\begin{figure}
\centering
\includegraphics[width=3.25in]{fig-emissions.png}
\caption{Mode shift diagram}\label{Femissions}
\end{figure}

@@

These usability problems were eventually solved by surveying other technical websites which organize and present large amounts of data. One website in particular, \url{Github.com}, was found to be well-liked by test users and has a similar hierarchical view of data: users can belong to organizations, and both organizations and users can create projects ("repositories" on Github) which may hold large numbers of files.

By adopting a project and user paradigm similar to that employed by Github, users were immediately more familiar with the site and had less to learn. The current site adheres to this so closely that test users often refer to the platform as "MatHub".

@@

\textbf{Map layer limitations.} The Mapbox GL mapping library allows lines, points, and polygons to be displayed on top of a base map. There seems to be a number of map elements beyond which performance becomes very slow; various techniques can be used such as aggregating elements at further-out zoom levels to get around this. Another option is to use Mapbox GL only for the base map, and to use the WebGL graphics libraries to draw large numbers of elements directly on top of the map. This is the technique used for the traffic animation, which can easily support tens of thousands of elements (all in motion!) simultaneously. For additional visualizations which have large numbers of grahical elements, more research will need to be done on layering WebGL elements on top of the base map.

@@

\bibitem[Strippgen, David (2016)]{R2}
Strippgen, David (2016) \textit{OTFVis: MATSim’s Open-Source Visualizer}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 225-–234.

\bibitem[Rieser, Marcel (2016)]{R3}
Rieser, Marcel (2016) \textit{Senozon Via}. In Andreas Horni, Kai Nagel, Kay W. Axhausen (Eds.): The Multi-Agent Transport Simulation MATSim: Ubiquity Press, pp. 219–-224.
